\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-31 metatheory-condensed.tex
linkage transformers

The third premise
$\goodTerm{\Gm,\TyTkg{\lsig}}{\seal}{\sub{A}{\SubstWeak{1}}}$
in rules \ruleref{l/add} and \ruleref{lsig/add} is
responsible for hiding W-type signatures in~$\lsig$.
Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{tyeq/pk/add})
packaging the types of all fields preceding the current field~$t$,
and $\seal$ is the term turning the tuple $\Tkg{\lkg}$ of type
$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new term of type $A$
that hides W-type signatures, if necessary.
When no hiding is needed, $\seal$ is the identity $\var{0}$ and $A$ is
$\TyTkg{\lsig}$.

\bigskip

A supplemental appendix gives a more complete example showing how families
are modeled and type-checked as linkages.

>>>>

The third premise
$\goodTerm{\Gm,\TyTkg{\lsig}}{\seal}{\sub{A}{\SubstWeak{1}}}$
in rules \ruleref{l/add} and \ruleref{lsig/add} is
responsible for hiding W-type signatures in~$\lsig$.
Here, $\TyTkg{\lsig}$ is a dependent tuple type (rule \ruleref{tyeq/pk/add})
packaging the types of all fields preceding the current field~$t$,
and $\seal$ is the term turning the tuple $\Tkg{\lkg}$ of type
$\TyTkg{\lsig}$ (rule \ruleref{tmeq/pk/add}) into a new term of type $A$
that hides W-type signatures behind $\cU$, if necessary.
When no hiding is needed, $\seal$ is $\var{0}$ (the identity) and $A$ is
$\TyTkg{\lsig}$.

When a family is closed and a complete linkage~$\lkg$ is formed, fields
of the family can be accessed by projecting them out of the tuple
$\Tkg{\lkg}$, as the last line of \cref{fig:stlc-linkage-typing} suggests.

\noindentparagraph{Linkage transformers.}

Inheritance and code reuse can already be expressed through projection
of fields out of linkages and their inclusion into new linkages.
To make common patterns of linkage manipulations more convenient,
\TT builds \emph{linkage transformers}, whose well-formedness judgments
have form $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$.
Linkage transformers are second-class; the only way to use them is to
apply them to linkages to form new linkages (\ruleref{tm/inh}).
%
Derived families can be modeled as linkage transformers.
%
Linkage transformers are inductively constructed from several introduction
rules.
\cref{fig:typing-selected} shows the beta rule of an exemplary transformer,
\ruleref{lxeq/ov/beta}.
It can read as follows: given a linkage of form $\LkgAdd{\lkg}{t_1}$,
applying the transformer $\InhOv{h}{t_2}$ overrides the linkage's last
field~$t_1$ with~$t_2$.
$\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$, for instance, can be used
in \cref{fig:stlc-linkage-typing} to form a derived family that
overrides $\wsig_{\texttt{tm}}$ with an extended
signature~$\wsig'_{\texttt{tm}}$.

\noindentparagraph{The complete formalization.}

The definitive version is formalized in a logical framework with
quotient inductive-inductive types~\cite{nordvall2010inductive,altkap2016},
available as a supplemental file in Agda syntax.
%
We also typeset a more complete set of rules than
\cref{fig:mltt-typing-selected,fig:fmltt-typing-selected} allow
and present it in a supplemental appendix.
%
Another supplemental appendix gives a larger-scale example than
\cref{fig:stlc-linkage-typing} showing how families are modeled and
type-checked using linkages and linkage transformers.
